% This is based on the LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
% See http://www.springer.com/computer/lncs/lncs+authors?SGWID=0-40209-0-0-0
% for the full guidelines.
%

\RequirePackage{amsmath}

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

%\usepackage{amsmath} % neprepisuji LLNCS symbol
\usepackage{amsfonts}
\usepackage[
backend=biber, 
style=numeric]{biblatex}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[caption=false]{subfig}
\usepackage{tikz}
\usepackage{url}
\usepackage{underscore}
\usepackage{pgffor}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{newunicodechar}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{arrayjobx}
\usepackage{tikz}
\usepackage{textcomp}
\usepackage{xifthen}
\usepackage{mleftright}
\usepackage{enumitem}
\usepackage{fancyvrb}
\usepackage{float}
\usepackage{alltt}
\usepackage{hhline}
\usepackage{environ}
\usepackage{adjustbox}
\usepackage{mathtools}
\usepackage[short]{optidef}
\usepackage{longtable}
\usepackage{lscape}
\usepackage{makecell}
\usepackage{colortbl}
\usepackage{listings}
\usepackage{wrapfig}
\usepackage{hyperref}
\usepackage{booktabs}

\usepackage{url}
\urldef{\mailsa}\path|{martin.beseda, lubomir.riha, alexandros.markopoulos, petr.strakos}@vsb.cz|

\pgfplotsset{compat=1.14}

\DeclareFieldFormat{labelnumberwidth}{#1\adddot}
\setlength{\biblabelsep}{5pt}

\DeclareCaptionLabelFormat{andtable}{#1~#2  \&  \tablename~\thetable}

\addbibresource{bibliography.bib}
\addbibresource{statistics.bib}
\addbibresource{feti.bib}
\addbibresource{mpi.bib}
\addbibresource{directSolvers.bib}
\addbibresource{pardiso.bib}
\addbibresource{readex.bib}
\addbibresource{others.bib}
\newcommand*{\alert}{\textcolor{red}}

\makeatletter
% mynobreakpar
\newcommand\mynobreakpar{\par\nobreak\@afterheading} 
%
\renewcommand*\env@matrix[1][*\c@MaxMatrixCols c]{%
  \hskip -\arraycolsep
  \let\@ifnextchar\new@ifnextchar
  \array{#1}}
\makeatother

\allowdisplaybreaks

\lstset{
    literate={~} {$\sim$}{1}
}

\begin{document}

\title{Performance Modeling of the HTFETI Solver Implementation in the ESPRESO Library}
%
\titlerunning{Performance Modeling the HTFETI Solver}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Martin Beseda\inst{a} \and Lubom\'{i}r \v{R}\'{i}ha\inst{a} \and Alexandros Markopoulos\inst{a} \and Petr Strako\v{s}\inst{a}}
%
\authorrunning{Performance Modeling the HTFETI Solver} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
%Craig Chambers, Kim B. Bruce, and Elisa Bertino}
%
\institute{$^a$IT4Innovations National Supercomputing Centre, VSB-Technical University of Ostrava, Ostrava, Czech Republic\\
%\email{martin.beseda@vsb.cz, lubomir.riha@vsb.cz, alexandros.markopoulos@vsb.cz}
\mailsa\\
\texttt{http://www.it4i.cz}
}
%\\ WWW home page:
%\texttt{http://users/\homedir iekeland/web/welcome.html}
%\and
%Universit\'{e} de Paris-Sud,
%Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
%F-91405 Orsay Cedex, France}

\maketitle              % typeset the title of the contribution

\begin{abstract}

The main objective of this paper is the development and implementation of the 
performance model for prediction of the runtime of the 
Total FETI and Hybrid Total FETI methods implemented in the ESPRESO library.

The final performance model consists of several partial ones, each created by 
a generalized linear regression, using R-language. These models were created from 
set of calibration measurements.

The final performance model is designed for prediction of the optimal settings 
of the solvers in a sense of optimal domain decomposition and optimal number of 
MPI processes and OpenMP threads. The user only provides the size of the problem 
to be solved.

Our tests show that the final model can predict optimal settings with an error 
smaller than 1s for a solver runtime that can vary between 3 and 50 seconds 
based on problem decomposition and parallelization setup.  

\keywords{performance model, ESPRESO, FETI, MPI, OpenMP}
\end{abstract}
%

\section{Motivation}
The motivation for the work presented in this paper is a fact that there exist several 
implementations of various FETI methods but it is very difficult for their users to 
setup them efficiently to minimize the runtime of the solver and to efficiently utilize the 
given hardware. 

Therefore this paper proposes a performance model for a massively parallel implementation of 
the Total FETI (TFETI) and Hybrid Total FETI (HTFETI) solvers in the 
ESPRESO library \cite{vriha2016massively}.


This model allows any inexperienced user to set-up both solvers in sense of 
(1) optimal domain decomposition and 
(2) optimal OpenMP and MPI parallelization, 
to minimize the solver runtime. 
Our model takes into account the parallel scalability of the solver however 
it is not able to predict the numerical scalability. In other words, model determines the minimal single iteration time, but it cannot predict how many iterations is needed 
to solve the problem because it is highly problem dependent.  


The model is composed of several partial models. Each of these estimates the run-time of a separate region. This enables easier customization of the model for future 
versions of ESPRESO as well as more detailed information of the solver behavior for 
different settings. 



\input{feti.tex}
%\input{numa.tex}

\section{Performance Modeling of the FETI Solvers }
\input{problemDesc.tex}
\input{modelling.tex}

\input{results.tex}
\input{conclusion.tex}

{\renewcommand{\markboth}[2]{}%
\printbibliography[heading=bibintoc]}
%\bibliographystyle{splncs03}
%\bibliography{References}

\appendix
\input{rSyntax.tex}
\input{partialModels.tex}




\end{document}
